A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2016-12-20 imaging_demo/ms.tex
Adhere to British spelling
From 0236a60 (mnras):
====================

In ms.tex,

(*) Remove comment on all arrays being heterogeneous
(*) Adhere to British spelling for colour, neighbour, centre, and
grey

The gridding kernel in the simplest case is given by the antenna aperture
illumination function, $W(\mathbf{r}-\mathbf{r}_a)$, which can
be specified either by a functional form or as a table of values against 
locations around the antennas. A nearest neighbor mapping from all antenna 
footprints to grid locations is created using an efficient k-d tree algorithm 
\citep{man99}. There is no restriction here that the aperture illumination 
function has to be identical across antennas.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure4}
  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
  \caption{Auto-correlation of weights of a square shaped antenna aperture
    in the $uv$ plane (left) and the corresponding directional antenna power 
    response on the sky (right) in coordinates specified by direction cosines. 
    The antenna auto-correlation weights are normalized to a sum
    of unity yielding a peak response of unity in the antenna's directional
    power pattern on the sky. The color scale for the directional power 
    pattern is logarithmic. The black circle indicates the sky horizon and
    values beyond it are not physical and hence ignored.}
  \label{fig:autocorr_wts_PB}
\end{figure}

@@

We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}. Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed gray) imaging. The two are found to match well. A magnified view shows that some differences at the level of $<0.5$\% exist in some regions. These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure6}
  % \includegraphics[width=\columnwidth]{MOFF_FX_synthesized_beam_slices}
  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line). The two appear almost identical. The inset provides a magnified view of differences in the synthesized beam slices between the two techniques at levels $\lesssim 0.5$\% relative to the peak. These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}
  \label{fig:MOFF-FX-psf-diff}
\end{figure}

The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging. The maximum difference appears near the center of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former. In other regions of the $uv$-plane, the differences are of the order of a few percent. The reason for these differences is described below.

@@

We study the effect of the differences in gridded weights on the image plane. Fig.~\ref{fig:image-psf-diff} shows the difference between the synthesized beams obtained with the two methods. A difference map between the two synthesized beams is shown on the left. The amplitude of the difference appears to be modulated by the directional power response of the antenna. On the right, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot). The antenna power pattern (red; to be read using the scale on the right) is plotted for reference.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure8}
  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
  \caption{Map of difference between the synthesized beams obtained with the 
    two methods (left) and radial statistics of the synthesized beams and 
    their differences (right). The radial variations of rms of the synthesized 
    beam (gray) and that of the rms of the difference (black) are shown as 
    percentage of the peak synthesized beam. Radially averaged directional 
    antenna power response in absolute scale is shown in red and is to be read 
    with the scale on the right side of the axis. The maximum difference is of 
    the order of a few percent. Amplitude of the rms of the difference is 
    modulated by the power pattern of the antenna.}
  \label{fig:image-psf-diff}
\end{figure}

@@

{\bf Heterogeneity in antenna arrays occurs primarily in two scenarios. The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as differences in gains between dipoles in a phased array. In practice, due to this reason almost all arrays are heterogeneous albeit by varying degrees. The second scenario is when the array is designed to be heterogeneous to gain specific advantages such as sampling a wider range of spatial frequencies and decoupling the sky brightness distribution from the antenna power pattern. This has motivated a number of heterogeneous arrays such as the Combined Array for Research in Millimeter-wave Astronomy (CARMA; \citealt{woo04,wri10}), the Atacama Large Millimeter/submillimeter Array (ALMA; \citealt{igu09}), and the Very Long Baseline Array (VLBA). The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example with large differences across antennas, making it closer to the second scenario.}

@@

In our simulations, the antenna layout, duration of observation, center frequency and bandwidth used are the same as in \S\ref{sec:sim}. Antennas have square apertures. Those of the first and second types, which have roughly equal numbers of antennas with similar layouts, have 1.1~m and 6.6~m on their sides respectively. $\mathcal{I}_\textrm{M}$ consists of point sources of random flux densities brighter than 10~Jy along westward and south-westward directions as seen in Fig.~\ref{fig:nonphysical-image}, which shows an optimally weighted image, $\mathcal{I}^\prime$, obtained with EPIC using the MOFF algorithm after averaging over the entire bandwidth and duration of simulated observation.

\begin{figure*}
\subfloat[][Optimally weighted EPIC image]{\label{fig:nonphysical-image}\includegraphics[width=0.5\linewidth]{figure10a.eps}}
\subfloat[][Estimates of normalized radial flux density profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation. 51 antennas in the inner core of the MWA layout were used with uniformly illuminated square apertures. Roughly half of them were squares with 1.1~m sides (first type) and the rest with 6.6~m (second type). Point sources were simulated along westward and south-westward directions with random flux densities brighter than 10~Jy. Black boxes shown around point sources are for purposes of collecting image statistics. (b)~Normalized flux density estimates, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, as a function of distance from image center. Bands indicate an approximate ``$1\sigma$'' uncertainty arising from sidelobe confusion and intrinsic randomness of electric fields simulated for point sources in the sky model. Gray and red bands correspond to optimal and erroneous cases respectively, where in the latter, all antennas were erroneously assumed to be identical and of the first type. The dashed and solid line styles correspond to point sources in the westward and south-westward directions respectively. In the optimal case, the normalized estimates are consistent with unity to within ``$1\sigma$'' in almost all regions of the image, whereas in the erroneous case, they are mis-estimated (underestimated) with high significance ($\gtrsim 5\sigma$) in most regions of the image. There is no significant difference in estimates between the two different directions of point sources studied.}
\label{fig:versatility}
\end{figure*}

@@

Fig.~\ref{fig:norm-flux-density} shows $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ including the uncertainty as a function of distance from the center of the image. The gray and red bands denote estimates from the optimal and the erroneous cases, respectively. The solid and dashed lines correspond to point sources in the south-westward and westward directions respectively. It can be clearly seen that the estimates in the optimal case in almost all parts of the image are consistent with unity to within $1\sigma$. In complete contrast, in the erroneous case, the estimated values appear to be underestimated and deviate from the ideal value of unity with very high significance ($\gtrsim 5\sigma$) in most parts of the image. The reason for underestimation is because nearly half the antennas are assumed to be smaller than they actually are and thus results in dividing by overestimated $\mathcal{W}_\textrm{eff}^\textrm{err}$. These trends do not seem to be affected by the placement of sources along westward and south-westward directions.

@@

Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
complex operations per frequency channel per cross-polarization per integration 
time-scale. Telescopes that fall to the left of the solid line indicate MOFF 
imaging is computationally more efficient than visibility-based imaging. All HERA 
layouts except HERA-19 and HERA-37 are in a region of parameter space 
where MOFF imaging holds the advantage. The dashed line showing future trajectory 
of HERA like systems will be clearly favoured by MOFF imaging. The dotted line is 
similarly a hypothetical trajectory for the MWA with more tiles added inside the 
same core diameter. The gray shaded area is for a projected LWA expansion and is 
also predominantly in the region favouring MOFF imaging. It is bounded by the LWA1 
and LWA-OV on the left and right respectively. The current (see 
Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a 
four-fold increase in number of elements over a 50 per cent increase in 
$b_\textrm{max}$ provide the bounds at the bottom and top respectively. Current 
instruments such as MWA and LOFAR lie in parameter space favouring 
visibility-based imaging.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure11}
  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
  \caption{Current and planned instruments in parameter space of
    number of complex {\bf multiplications and additions}. The solid line
    is the boundary at which the number of operations with MOFF and 
    visibility-based imaging are equal. MOFF imaging is more efficient for 
    telescopes occupying the left of this line and vice versa. CHIME, HIRAX, 
    SKA1-LC, SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie 
    in the parameter space favoured by MOFF imaging. The dashed line shows the 
    projected trajectory of hypothetical expanded HERA layouts. The dotted line 
    similarly shows hypothetical expanded MWA layouts with more tiles added in 
    the same core. The gray shaded area denotes the projected trajectory of the 
    LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
    (bottom) and a four-fold increase in the number of elements within a 50 per 
    cent increase in the core size (top). Current instruments such as MWA and 
    LOFAR fall in a region favoured by visibility-based imaging.}
  \label{fig:parameter-space-computations-instruments}
\end{figure}

We now consider antenna array layouts described by three essential quantities in radio interferometry, namely, maximum baseline length, number of antennas, and the size of each antenna. Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where the ratio of the number of computations required with visibility-based imaging relative to MOFF imaging is unity. The different colored lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$). Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure12}
  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2}
  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2.eps}
  \caption{Current and planned instruments in parameter space of baseline length and number of antennas. Lines of different colors denote different classes of antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$). Lines of each color denote the boundary to the left of which MOFF algorithm is favoured for the corresponding antenna size. These lines shift rightward with increasing antenna size. The different antennas are color coded by roughly the class of antenna size they fall into. Thus symbols of one color falling to the left of a line of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa. For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.}
  \label{fig:parameter-space-bll-nant-instruments}
\end{figure}

@@

The different antennas are color coded by roughly the class of antenna size they fall into. Thus symbols of one color that lie to the left of same colored line indicate MOFF imaging is favoured for those telescopes and vice versa. For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37. A majority of the next-generation radio telescopes, namely, HERA-331 and its future expanded versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF imaging will be desirable. LWA1 and LWA-OV are already very close to the dividing line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in number of antennas within a core diameter that is 1 and 1.5 times the current size of 100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational regime favoured by MOFF. For a fixed baseline length, regions favouring the MOFF algorithm tend to be towards large $\Nant$ indicating large-N dense array layouts with smaller antenna elements are best suited for deploying EPIC.

>>>>

The gridding kernel in the simplest case is given by the antenna aperture
illumination function, $W(\mathbf{r}-\mathbf{r}_a)$, which can
be specified either by a functional form or as a table of values against 
locations around the antennas. A nearest neighbour mapping from all antenna 
footprints to grid locations is created using an efficient k-d tree algorithm 
\citep{man99}. There is no restriction here that the aperture illumination 
function has to be identical across antennas.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figure4}
  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
  \caption{Auto-correlation of weights of a square shaped antenna aperture
    in the $uv$ plane (left) and the corresponding directional antenna power 
    response on the sky (right) in coordinates specified by direction cosines. 
    The antenna auto-correlation weights are normalized to a sum
    of unity yielding a peak response of unity in the antenna's directional
    power pattern on the sky. The colour scale for the directional power 
    pattern is logarithmic. The black circle indicates the sky horizon and
    values beyond it are not physical and hence ignored.}
  \label{fig:autocorr_wts_PB}
\end{figure}

@@

We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}. Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed grey) imaging. The two are found to match well. A magnified view shows that some differences at the level of $<0.5$\% exist in some regions. These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure6}
  % \includegraphics[width=\columnwidth]{MOFF_FX_synthesized_beam_slices}
  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed grey line). The two appear almost identical. The inset provides a magnified view of differences in the synthesized beam slices between the two techniques at levels $\lesssim 0.5$\% relative to the peak. These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}
  \label{fig:MOFF-FX-psf-diff}
\end{figure}

The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging. The maximum difference appears near the centre of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former. In other regions of the $uv$-plane, the differences are of the order of a few percent. The reason for these differences is described below.

@@

We study the effect of the differences in gridded weights on the image plane. Fig.~\ref{fig:image-psf-diff} shows the difference between the synthesized beams obtained with the two methods. A difference map between the two synthesized beams is shown on the left. The amplitude of the difference appears to be modulated by the directional power response of the antenna. On the right, in radial bins, the rms of the synthesized beam (grey) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot). The antenna power pattern (red; to be read using the scale on the right) is plotted for reference.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure8}
  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
  \caption{Map of difference between the synthesized beams obtained with the 
    two methods (left) and radial statistics of the synthesized beams and 
    their differences (right). The radial variations of rms of the synthesized 
    beam (grey) and that of the rms of the difference (black) are shown as 
    percentage of the peak synthesized beam. Radially averaged directional 
    antenna power response in absolute scale is shown in red and is to be read 
    with the scale on the right side of the axis. The maximum difference is of 
    the order of a few percent. Amplitude of the rms of the difference is 
    modulated by the power pattern of the antenna.}
  \label{fig:image-psf-diff}
\end{figure}

@@

{\bf Heterogeneity in antenna arrays occurs primarily in two scenarios. The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as differences in gains between dipoles in a phased array. % In practice, due to this reason almost all arrays are heterogeneous albeit by varying degrees. 
  The second scenario is when the array is designed to be heterogeneous to gain specific advantages such as sampling a wider range of spatial frequencies and decoupling the sky brightness distribution from the antenna power pattern. This has motivated a number of heterogeneous arrays such as the Combined Array for Research in Millimeter-wave Astronomy (CARMA; \citealt{woo04,wri10}), the Atacama Large Millimeter/submillimeter Array (ALMA; \citealt{igu09}), and the Very Long Baseline Array (VLBA). The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example with large differences across antennas, making it closer to the second scenario.}

@@

In our simulations, the antenna layout, duration of observation, centre frequency and bandwidth used are the same as in \S\ref{sec:sim}. Antennas have square apertures. Those of the first and second types, which have roughly equal numbers of antennas with similar layouts, have 1.1~m and 6.6~m on their sides respectively. $\mathcal{I}_\textrm{M}$ consists of point sources of random flux densities brighter than 10~Jy along westward and south-westward directions as seen in Fig.~\ref{fig:nonphysical-image}, which shows an optimally weighted image, $\mathcal{I}^\prime$, obtained with EPIC using the MOFF algorithm after averaging over the entire bandwidth and duration of simulated observation.

\begin{figure*}
\subfloat[][Optimally weighted EPIC image]{\label{fig:nonphysical-image}\includegraphics[width=0.5\linewidth]{figure10a.eps}}
\subfloat[][Estimates of normalized radial flux density profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation. 51 antennas in the inner core of the MWA layout were used with uniformly illuminated square apertures. Roughly half of them were squares with 1.1~m sides (first type) and the rest with 6.6~m (second type). Point sources were simulated along westward and south-westward directions with random flux densities brighter than 10~Jy. Black boxes shown around point sources are for purposes of collecting image statistics. (b)~Normalized flux density estimates, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, as a function of distance from image centre. Bands indicate an approximate ``$1\sigma$'' uncertainty arising from sidelobe confusion and intrinsic randomness of electric fields simulated for point sources in the sky model. Grey and red bands correspond to optimal and erroneous cases respectively, where in the latter, all antennas were erroneously assumed to be identical and of the first type. The dashed and solid line styles correspond to point sources in the westward and south-westward directions respectively. In the optimal case, the normalized estimates are consistent with unity to within ``$1\sigma$'' in almost all regions of the image, whereas in the erroneous case, they are mis-estimated (underestimated) with high significance ($\gtrsim 5\sigma$) in most regions of the image. There is no significant difference in estimates between the two different directions of point sources studied.}
\label{fig:versatility}
\end{figure*}

@@

Fig.~\ref{fig:norm-flux-density} shows $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ including the uncertainty as a function of distance from the centre of the image. The grey and red bands denote estimates from the optimal and the erroneous cases, respectively. The solid and dashed lines correspond to point sources in the south-westward and westward directions respectively. It can be clearly seen that the estimates in the optimal case in almost all parts of the image are consistent with unity to within $1\sigma$. In complete contrast, in the erroneous case, the estimated values appear to be underestimated and deviate from the ideal value of unity with very high significance ($\gtrsim 5\sigma$) in most parts of the image. The reason for underestimation is because nearly half the antennas are assumed to be smaller than they actually are and thus results in dividing by overestimated $\mathcal{W}_\textrm{eff}^\textrm{err}$. These trends do not seem to be affected by the placement of sources along westward and south-westward directions.

@@

Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
complex operations per frequency channel per cross-polarization per integration 
time-scale. Telescopes that fall to the left of the solid line indicate MOFF 
imaging is computationally more efficient than visibility-based imaging. All HERA 
layouts except HERA-19 and HERA-37 are in a region of parameter space 
where MOFF imaging holds the advantage. The dashed line showing future trajectory 
of HERA like systems will be clearly favoured by MOFF imaging. The dotted line is 
similarly a hypothetical trajectory for the MWA with more tiles added inside the 
same core diameter. The grey shaded area is for a projected LWA expansion and is 
also predominantly in the region favouring MOFF imaging. It is bounded by the LWA1 
and LWA-OV on the left and right respectively. The current (see 
Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a 
four-fold increase in number of elements over a 50 per cent increase in 
$b_\textrm{max}$ provide the bounds at the bottom and top respectively. Current 
instruments such as MWA and LOFAR lie in parameter space favouring 
visibility-based imaging.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure11}
  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
  \caption{Current and planned instruments in parameter space of
    number of complex {\bf multiplications and additions}. The solid line
    is the boundary at which the number of operations with MOFF and 
    visibility-based imaging are equal. MOFF imaging is more efficient for 
    telescopes occupying the left of this line and vice versa. CHIME, HIRAX, 
    SKA1-LC, SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie 
    in the parameter space favoured by MOFF imaging. The dashed line shows the 
    projected trajectory of hypothetical expanded HERA layouts. The dotted line 
    similarly shows hypothetical expanded MWA layouts with more tiles added in 
    the same core. The grey shaded area denotes the projected trajectory of the 
    LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
    (bottom) and a four-fold increase in the number of elements within a 50 per 
    cent increase in the core size (top). Current instruments such as MWA and 
    LOFAR fall in a region favoured by visibility-based imaging.}
  \label{fig:parameter-space-computations-instruments}
\end{figure}

We now consider antenna array layouts described by three essential quantities in radio interferometry, namely, maximum baseline length, number of antennas, and the size of each antenna. Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where the ratio of the number of computations required with visibility-based imaging relative to MOFF imaging is unity. The different coloured lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, grey - 5900~m$^2$). Dashed lines of each colour denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure12}
  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2}
  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2.eps}
  \caption{Current and planned instruments in parameter space of baseline length and number of antennas. Lines of different colours denote different classes of antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, grey - 5900~m$^2$). Lines of each colour denote the boundary to the left of which MOFF algorithm is favoured for the corresponding antenna size. These lines shift rightward with increasing antenna size. The different antennas are colour coded by roughly the class of antenna size they fall into. Thus symbols of one colour falling to the left of a line of the same colour indicate MOFF imaging is advantageous for those telescopes and vice versa. For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.}
  \label{fig:parameter-space-bll-nant-instruments}
\end{figure}

@@

The different antennas are colour coded by roughly the class of antenna size they fall into. Thus symbols of one colour that lie to the left of same coloured line indicate MOFF imaging is favoured for those telescopes and vice versa. For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37. A majority of the next-generation radio telescopes, namely, HERA-331 and its future expanded versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF imaging will be desirable. LWA1 and LWA-OV are already very close to the dividing line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in number of antennas within a core diameter that is 1 and 1.5 times the current size of 100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational regime favoured by MOFF. For a fixed baseline length, regions favouring the MOFF algorithm tend to be towards large $\Nant$ indicating large-N dense array layouts with smaller antenna elements are best suited for deploying EPIC.
